Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

Unfortunately, none of these fragments can accommodate the vast majority of
quantified formulas that correspond to recursive function definitions: The
essentially uninterpreted fragment does not allow the argument of a
recursive function to be used inside a complex term on the right-hand side,
whereas the finite model finding techniques of Reynolds et al.\ are not
applicable for functions over infinite domains such as the integers or
algebraic datatypes. A simple example where both approaches fail is
\begin{equation}\label{eq:pii}
\forall x : {\typeint}.\;\, \lite\bigl(
x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)
\end{equation}
This state of affairs is highly unsatisfactory, given the frequence of
recursive definitions in practice and the impending addition of a dedicated
command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.

@@

The Leon system~\cite{blanc2013overview} implements a procedure that can both
produces proofs and counterexamples for properties of terminating functions
written in a subset of Scala. Leon is based on an SMT solver. It avoids
quantifiers altogether by unfolding recursive definitions up to a certain
depth, which is increased on a per-need basis.
The approach described in this paper works in an analagous manner, 
where instead the SMT solver is invoked only once 
and where quantifier instantiation amounts to function unfolding. It would be
worth investigating how existing approaches for algorithms based on function
unfolding can inform approaches for dedicated quantifier instantiation
techniques for function definitions, and vice versa.

>>>>

Unfortunately, none of these fragments can accommodate the vast majority of
quantified formulas that correspond to recursive function definitions: The
essentially uninterpreted fragment does not allow the argument of a
recursive function to be used inside a complex term on the right-hand side,
whereas the finite model finding techniques of Reynolds et al.\ are not
applicable for functions over infinite domains such as the integers or
algebraic datatypes. A simple example where both approaches fail is
\begin{equation}\label{eq:pii}
\forall x : {\typeint}.\;\, \lite\bigl(
x \leq 0,\allowbreak\; \con{p}( x ) \teq 1,\allowbreak\; \con{p}( x ) \teq 2 * \con{p}( x - 1 ) \bigr)
\end{equation}
This state of affairs is highly unsatisfactory, given the frequency of
recursive definitions in practice and the impending addition of a dedicated
command, \texttt{define-fun-rec}, to the SMT-LIB standard \cite{smtlib25}.

@@

The Leon system~\cite{blanc2013overview} implements a procedure that can both
produces proofs and counterexamples for properties of terminating functions
written in a subset of Scala. Leon is based on an SMT solver. It avoids
quantifiers altogether by unfolding recursive definitions up to a certain
depth, which is increased on a per-need basis.
The approach described in this paper works in an analogous manner, 
where instead the SMT solver is invoked only once 
and where quantifier instantiation amounts to function unfolding. It would be
worth investigating how existing approaches for algorithms based on function
unfolding can inform approaches for dedicated quantifier instantiation
techniques for function definitions, and vice versa.
